Supporting Information:\\ NMRlipids IV: Headgroup \& glycerol backbone structures, and cation binding in bilayers with PS lipids

### 2018-12-10 Manuscript/manuscriptPS.tex
Update simulation details

\begin{table*}[htb]
%\begin{sidewaystable*}[!p]
\centering
\caption{The list of MD simulations of pure PS bilayers without additional salt.
  Simulation details are given in the supplementary information
    %   The lipid force fields named as in our previous work~\cite{botan15}.
}\label{PSsystems}
%begin{minipage}[t]{\textwidth}
\begin{tabular}{l c r r r r r c c}
  %\hline
  % some footnotes are not visible in typeset-MS (pdf)
 lipid/counter-ions & force field for lipids / ions & \footnote{Number of lipid molecules with largest mole fraction}N$_{\rm l}$   &  \footnote{Number of water molecules}N$_{\rm w}$  & \footnote{Number of additional cations}N$_{\rm c}$  & \footnote{Simulation temperature}T (K)  & \footnote{Total simulation time}t$_{{\rm sim}}$(ns) & \footnote{Time used for analysis}t$_{{\rm anal}}$ (ns) &   \footnote{Reference for simulation files}files\\
  \hline
    DOPS/Na$^+$  & CHARMM36 \cite{venable13}       & 128 & 4480 & 0  & 303  & 500 & 100 & \cite{charmm36DOPS303K} \\
    DOPS/Na$^+$  & CHARMM36ua \cite{??} \todoi{Correct citation for CHARMMua DOPS}   & 128 & 4480 & 0  & 303  & 500 & 100 & \cite{charmm36uaDOPS303K} \\
    DOPS/Na$^+$  & Slipids \cite{jambeck13}        & 128 	& 4480  & 0  & 303  & 500 & 100 & \cite{slipidsDOPS303K} \\
    DOPS/Na$^+$  & Slipids \cite{jambeck13}        & 288 	& 11232 & 0  & 303  & 200 & 100 & \cite{slipidsDOPSfiles} \\
    DOPS/Na$^+$  & Berger \cite{mukhopadhyay04}    & 128  & 4480  & 0  & 303  & 500 & 100 & \cite{bergerDOPS303K} \\
    DOPS/Na$^+$  & GROMOS-CKPM \cite{??} \todoi{Correct citation(s) for CKP.} & 128 & 4480 & 0  & 303  & 500 & 100 & \cite{ckp1DOPS303K} \\
    DOPS/Na$^+$  & GROMOS-CKP  \cite{??} \todoi{Correct citation(s) for CKP.} & 128 & 4480 & 0  & 303  & 500 & 100 & \cite{ckp2DOPS303K} \\
    DOPS/Na$^+$  & lipid17 \cite{gould18} / JC  \cite{joung08}    & 128    & 4480   & 0   & 303  & 600 & 100 & \cite{lipid17DOPSjcions} \\
    DOPS/Na$^+$  & lipid17 \cite{gould18} / ff99 \cite{aqvist90}  & 128    & 4480   & 0   & 303  & 600 & 100 & \cite{lipid17DOPSff99ions} \\
    \hline
    POPS/Na$^+$  & CHARMM36 \cite{venable13} & 128 & 4480 & 0  & 298  & 500 & 100 & \cite{charmm36POPS298K} \\
    POPS/K$^+$   & CHARMM36 \cite{venable13} & 128 & 4480 & 0  & 298  & 500 & 100 & \cite{charmm36POPS298Kpotassium} \\
    POPS/Na$^+$  & CHARMM36ua \cite{??} \todoi{Correct citation for CHARMMua DOPS} & 128 & 4480 & 0  & 298  & 500 & 100 & \cite{charmm36uaPOPS298K} \\
    POPS/Na$^+$  & Slipids \cite{jambeck13}  & 128 & 4480 & 0  & 298  & 500 & 100 & \cite{slipidsPOPS298K} \\
    POPS/Na$^+$  & Berger \cite{??}          & 128 & 4480 & 0  & 298  & 500 & 100 & \cite{bergerPOPS298K} \\
    POPS/Na$^+$  & MacRog \cite{maciejewski14}  & 128 & 4480 & 0  & 298  & 500 & 100  & \cite{macrogPOPS298Kcorrect} \\
    POPS/K$^+$   & MacRog \cite{maciejewski14}  & 128 & 4480 & 0  & 298  & 200 & 150 & \cite{macrogPOPS298KwithK} \\
    POPS/Na$^+$  & GROMOS-CKPM \cite{??} \todoi{Correct citation(s) for CKP.} & 128 & 4480 & 0  & 298  & 500 & 100 & \cite{ckp1POPS303K} \\
    POPS/Na$^+$  & GROMOS-CKP \cite{??} \todoi{Correct citation(s) for CKP.}  & 128 & 4480 & 0  & 298  & 500 & 100 & \cite{ckp2POPS303K} \\
    POPS/Na$^+$  & lipid17  \cite{gould18} / JC  \cite{joung08}   & 128    & 4480   & 0   & 298  & 600 & 100 & \cite{lipid17POPSjcions} \\
    POPS/Na$^+$  & lipid17 \cite{gould18} / ff99 \cite{aqvist90}  & 128    & 4480   & 0   & 298  & 600 & 100 & \cite{lipid17POPSff99ions} \\
    \end{tabular}
%\end{minipage}
%\end{sidewaystable*} 
\end{table*}

@@

    POPC:POPS (5:1)/K$^+$  & CHARMM36 \cite{klauda10,venable13} &0 & 0  & 110:22 & 4935 & 0  & 298  & 100 & 100 \todoi{Equilibration?} & \cite{charmm36pops+83popcT298K}  \\

>>>>

\begin{table*}[htb]
%\begin{sidewaystable*}[!p]
\centering
\caption{The list of MD simulations of pure PS bilayers without additional salt.
  Simulation details are given in the supplementary information
    %   The lipid force fields named as in our previous work~\cite{botan15}.
}\label{PSsystems}
%begin{minipage}[t]{\textwidth}
\begin{tabular}{l c r r r r r c c}
  %\hline
  % some footnotes are not visible in typeset-MS (pdf)
 lipid/counter-ions & force field for lipids / ions & \footnote{Number of lipid molecules with largest mole fraction}N$_{\rm l}$   &  \footnote{Number of water molecules}N$_{\rm w}$  & \footnote{Number of additional cations}N$_{\rm c}$  & \footnote{Simulation temperature}T (K)  & \footnote{Total simulation time}t$_{{\rm sim}}$(ns) & \footnote{Time used for analysis}t$_{{\rm anal}}$ (ns) &   \footnote{Reference for simulation files}files\\
    \hline
    POPS/Na$^+$  & CHARMM36 \cite{venable13} & 128 & 4480 & 0  & 298  & 500 & 100 & \cite{charmm36POPS298K} \\
    POPS/K$^+$   & CHARMM36 \cite{venable13} & 128 & 4480 & 0  & 298  & 500 & 100 & \cite{charmm36POPS298Kpotassium} \\
    POPS/Na$^+$  & CHARMM36ua \cite{??} \todoi{Correct citation for CHARMMua DOPS} & 128 & 4480 & 0  & 298  & 500 & 100 & \cite{charmm36uaPOPS298K} \\
    POPS/Na$^+$  & MacRog \cite{maciejewski14}  & 128 & 4480 & 0  & 298  & 500 & 100  & \cite{macrogPOPS298Kcorrect} \\
    POPS/K$^+$   & MacRog \cite{maciejewski14}  & 128 & 4480 & 0  & 298  & 200 & 150 & \cite{macrogPOPS298KwithK} \\
    POPS/Na$^+$  & lipid17  \cite{gould18} / JC  \cite{joung08}   & 128    & 4480   & 0   & 298  & 600 & 100 & \cite{lipid17POPSjcions} \\
    POPS/Na$^+$  & lipid17 \cite{gould18} / ff99 \cite{aqvist90}  & 128    & 4480   & 0   & 298  & 600 & 100 & \cite{lipid17POPSff99ions} \\
    POPS/Na$^+$  & Berger \cite{mukhopadhyay04,??}          & 128 & 4480 & 0  & 298  & 500 & 100 & \cite{bergerPOPS298K} \\
    POPS/Na$^+$  & Slipids \cite{jambeck13}  & 128 & 4480 & 0  & 298  & 500 & 100 & \cite{slipidsPOPS298K} \\
    POPS/Na$^+$  & GROMOS-CKPM \cite{??} \todoi{Correct citation(s) for CKP.} & 128 & 4480 & 0  & 298  & 500 & 100 & \cite{ckp1POPS303K} \\
    POPS/Na$^+$  & GROMOS-CKP \cite{??} \todoi{Correct citation(s) for CKP.}  & 128 & 4480 & 0  & 298  & 500 & 100 & \cite{ckp2POPS303K} \\
  \hline
    DOPS/Na$^+$  & CHARMM36 \cite{venable13}       & 128 & 4480 & 0  & 303  & 500 & 100 & \cite{charmm36DOPS303K} \\
    DOPS/Na$^+$  & CHARMM36ua \cite{??} \todoi{Correct citation for CHARMMua DOPS}   & 128 & 4480 & 0  & 303  & 500 & 100 & \cite{charmm36uaDOPS303K} \\
    DOPS/Na$^+$  & lipid17 \cite{gould18} / JC  \cite{joung08}    & 128    & 4480   & 0   & 303  & 600 & 100 & \cite{lipid17DOPSjcions} \\
    DOPS/Na$^+$  & lipid17 \cite{gould18} / ff99 \cite{aqvist90}  & 128    & 4480   & 0   & 303  & 600 & 100 & \cite{lipid17DOPSff99ions} \\
    DOPS/Na$^+$  & Berger \cite{mukhopadhyay04,??}    & 128  & 4480  & 0  & 303  & 500 & 100 & \cite{bergerDOPS303K} \\
    DOPS/Na$^+$  & Slipids \cite{jambeck13}        & 128 	& 4480  & 0  & 303  & 500 & 100 & \cite{slipidsDOPS303K} \\
    DOPS/Na$^+$  & Slipids \cite{jambeck13}        & 288 	& 11232 & 0  & 303  & 200 & 100 & \cite{slipidsDOPSfiles} \\
    DOPS/Na$^+$  & GROMOS-CKPM \cite{??} \todoi{Correct citation(s) for CKP.} & 128 & 4480 & 0  & 303  & 500 & 100 & \cite{ckp1DOPS303K} \\
    DOPS/Na$^+$  & GROMOS-CKP  \cite{??} \todoi{Correct citation(s) for CKP.} & 128 & 4480 & 0  & 303  & 500 & 100 & \cite{ckp2DOPS303K} \\
    \end{tabular}
%\end{minipage}
%\end{sidewaystable*} 
\end{table*}

@@

%    POPC:POPS (5:1)/K$^+$  & CHARMM36 \cite{klauda10,venable13} &0 & 0  & 110:22 & 4935 & 0  & 298  & 100 & 100 \todoi{Equilibration?} & \cite{charmm36pops+83popcT298K}  \\

### 2018-12-10 Manuscript/manuscriptPSsuppl.tex
Update simulation details

\subsection{CHARMM36}
\todo{To be written by Piggot, Madsen and Ollila}

\subsection{CHARMM36ua}
\todo{To be written by Piggot}

@@

\subsection{Berger}
\todo{To be written by Piggot and Ollila}
Simulations with excess sodium were taken directly from Ref. \citenum{jurkiewicz12} and
simulations with calcium directly from \citenum{melcrova16}.
Simulation of POPC at 310 K was taken directly from Ref. \citenum{ollila07a}.

@@

\subsection{Lipid17}
\todo{To be written by Kav, Miettinen and Meclr.}

\subsection{MacRog}
\todo{To be written by Javanainen and Piggot}

>>>>

\subsection{CHARMM36}

\noindent
{\it POPC bilayer}
\todo{By Fernando Favela at http://dx.doi.org/10.5281/zenodo.16830} \\
% It should be noted that pure POPC are ran with Gromacs 4, but the results with
% PS are ran with Gromacs 5. This has significant effect on acyl chains:
% https://github.com/NMRLipids/NmrLipidsCholXray/issues/4
% The effect for headgroups can be estimated by comparing these two data sets:
% https://github.com/NMRLipids/MATCH/blob/master/Data/Lipid_Bilayers/POPC/T303K/MODEL_CHARMM36_GROMACS4.5/Headgroup_Glycerol_Order_Parameters_Simulation.dat
% https://github.com/NMRLipids/MATCH/blob/master/Data/Lipid_Bilayers/POPC/T303K/MODEL_CHARMM36_GROMACS5.1.2/Headgroup_Glycerol_Order_Parameters_Simulation.dat}

\noindent
{\it DOPS and POPS bilayers.} \\
The data in Table \ref{PSsystems} and Refs \citenum{charmm36DOPS303K,charmm36POPS298K,charmm36POPS298Kpotassium}
\todo{T. Piggot, please write simulation details} \\

\noindent
{\it POPC:POPS mixtures without additional ions}\\
Simulations of POPC:POPS (5:1) mixture containing total 132 (110:22) lipids
with potassium and sodium countertions were ran using CHARMM36 force field.
The data are available from Refs. \citenum{charmm36pops+83popcT298Kpiggot}
and \citenum{charmm36pops+83popcT298KpiggotSODIUM}, respectively\todo{T. Piggot, please write simulation details}.

Simulations of POPC:POPS (5:1) ans POPC:POPS (1:1) mixtures containing total 300 lipids
with potassium countertions were ran using CHARMM36 force field.
The data are available from Refs. \citenum{??} \todo{J. Madsen, please write simulation details}. \\

\noindent
{\it POPC:POPS (5:1) mixtures with additional monovalent ions} \\
POPC:POPS (5:1) mixtures containing total 132 (110:22) lipids
were simulated with the additional potassium and sodium countertion concentrations
of $\sim$450 mM and 890 mM using CHARMM36 force field at 298 K. The data, force field and
input files generated using parameters given by CHARMM-GUI \cite{lee16,jo18} are
available from Refs. \citenum{charmm36pops+83popcT298Kwith450mMK,charmm36pops+83popcT298Kwith890mMK,charmm36pops+83popcT298Kwith440mMNa,charmm36pops+83popcT298Kwith890mMNa}. \\

\noindent
{\it POPC:POPS (5:1) mixtures with additional calcium} \\
POPC:POPS (5:1) mixtures containing total 300 (250:50) lipids
were simulated with the additional calcium concentrations
\todo{J. Madsen, please write simulation details}. \\

\subsection{CHARMM36ua}
\noindent
{\it DOPS and POPS bilayers.} \\
The data in Table \ref{PSsystems} and Refs \citenum{charmm36uaDOPS303K,charmm36uaPOPS298K}
\todo{T. Piggot, please write simulation details} \\

\subsection{MacRog}
\noindent
{\it POPC bilayers.} \\

\noindent
{\it POPS bilayers.} \\
The data in Table \ref{PSsystems} and Refs \citenum{macrogPOPS298Kcorrect,macrogPOPS298KwithK}
\todo{T. Piggot and M. Javanainen, please write simulation details} \\

\noindent
{\it POPC:POPS (5:1) mixtures with additional potassium ions} \\
The data in Table \ref{mixedIONsystems} and Refs \citenum{POPCpopsMACROG,POPCpopsMACROGwithK}
\todo{M. Javanainen, please write simulation details} \\

\subsection{Lipid17}
{\it DOPS and POPS bilayers.} \\
The data in Table \ref{PSsystems} and Refs \citenum{lipid17POPSjcions,lipid17POPSff99ions,lipid17DOPSjcions,lipid17DOPSff99ions}
\todo{B. Kav and M. Miettinen, please write simulation details} \\

\noindent
{\it POPC:POPS (5:1) mixtures with additional potassium and sodium ions} \\
The data in Table \ref{mixedIONsystems} and Refs \citenum{POPCpopsLIPID17withKCI,POPCpopsLIPID17withK,POPCpopsLIPID17withNaCI,POPCpopsLIPID17withNa}
\todo{B. Kav and M. Miettinen, please write simulation details} \\

\noindent
{\it POPC:POPS (5:1) mixtures with additional calcium} \\
The data in Table \ref{mixedIONsystems} and Refs \citenum{lipid17_cacl_series}
\todo{J. Melcr, please write simulation details} \\

\subsection{Berger}
\noindent
{\it POPC bilayers.} \\
Previously published simulation of POPC at 310 K \cite{ollila07a}, available from Ref. \cite{bergerPOPC310K},
was used.

\noindent
{\it DOPS and POPS bilayers.} \\
The data in Table \ref{PSsystems} and Refs \citenum{bergerPOPS298K,bergerDOPS303K}
\todo{T. Piggot, please write simulation details} \\

\noindent
{\it POPC:POPS (4:1) mixtures with additional sodium and calcium ions} \\
Simulations with excess sodium were taken directly from Ref. \citenum{jurkiewicz12} and
simulations with calcium directly from \citenum{melcrova16}.

@@

@@
